---
title: "Homework 1"
author: "Jyoti Ankam"
date: "09-16-2018"
output: github_document
---

## Problem 1

The code chunk below creates a data frame for the following attributes, random_sample, logic_vec, char_vec and fact_vec. 

```{r}
library(tidyverse)
 
set.seed(5)

prob1_df = tibble(
  random_sample = runif(10, min = 0, max = 5),
  logic_vec = random_sample > 2,
  char_vec = c("Bear", "Deer", "Bat", "Hamster", "Parrot", "Wolf", "Owl", "Penguin", "Ibis", "Squirrel"),
  fact_vec = factor(c("Mammal", "Mammal", "Mammal", "Mammal", "Bird", "Mammal", "Bird", "Bird", "Bird", "Mammal"))
)

```

Obtaining means of random_sample, logic_vec, char_vec and fact_vec. 

```{r}
mean(prob1_df$random_sample)

```

The mean of variable random_sample is 2.64, which means it works.

```{r}
mean(prob1_df$logic_vec)

```
The mean of variable logic_vec is 0.6, which means it works.

```{r}
mean(prob1_df$char_vec)

```
The mean of variable char_vec is not obtainable, because the argument is not numeric or logical.


```{r}
mean(prob1_df$fact_vec)
```
The mean of variable fact_vec is not obtainable, because the argument is not numeric or logical.

Writing a code chunk that applies the as.numeric function to the logical, character, and factor variables:

```{r, results="hide"}
as.numeric(prob1_df$logic_vec)
as.numeric(prob1_df$char_vec)
as.numeric(prob1_df$fact_vec)
```
 
We can see that as.numeric for character variable did not work and we received a warning message that it is NA (NAs introduced by coercion)

In contrast, we could convert factor variable into numeric easily because they were assigned by R in a systematic order.


In a second code chunk, we are converting our character variable from character to factor to numeric; similarly, convert your factor variable from factor to character to numeric:

```{r}

fact_convert = as.factor(prob1_df$char_vec)

as.numeric(fact_convert)

char_convert = as.character(prob1_df$fact_vec)

as.numeric(char_convert)

```
Here, we converted the character variable from character to factor to numeric and numbers were assigned to the factors; 


Similarly, when we convert the factor variable from factor to character to numeric, we receive a warning message with NAs.


## Problem 2:
Creating a data frame comprising variables x, y, log_vec, numeric_vec and fac_vec.
```{r}
set.seed(4)
prob2_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  log_vec = x + y > 0,
  numeric_vec = as.numeric(log_vec),
  fac_vec = as.factor(log_vec)
)
```

This is the inline R code:


The size of this dataset is `r nrow(prob2_df)` and the mean and median of variable X are `r mean(prob2_df$x)` and `r median(prob2_df$x)`. The proportion of cases for which the logical vector is TRUE is `r (sum(prob2_df$log_vec)/nrow(prob2_df))`

Making a scatterplot of y vs x and saving the first scatterplot:

```{r}

library(ggplot2)

ggplot(prob2_df, aes(x = x, y = y, color = log_vec)) + geom_point()

ggsave("scatter1.pdf")

```

As we can see from the plots, ccatterplots for variables log_vec and fac_vec have only two option of TRUE (color coded Blue) and FALSE (color coded Pink) while variable numeric_vec has been color coded in terms of shades of blue depending the range of values from the minimum to the maximum. 

```{r}

ggplot(prob2_df, aes(x = x, y = y, color = numeric_vec)) + geom_point()

```

```{r}

ggplot(prob2_df, aes(x = x, y = y, color = fac_vec)) + geom_point()

```


